---
title: "Phan tich hoi tu GRDP/nguoi"
subtitle: "ca nuoc - 63 tinh/thanh"
author: "Vo Thanh Danh"
output:
  html_document:
    code_download: true
    df_print: paged
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
    number_sections: true
    code_folding: "show"
    theme: "cosmo"
    highlight: "monochrome"
  pdf_document: default
  word_document: default
---


<style>
h1.title {font-size: 18pt; color: DarkBlue;} 
body, h1, h2, h3, h4 {font-family: "Palatino", serif;}
body {font-size: 12pt;}
/* Headers */
h1,h2,h3,h4,h5,h6{font-size: 14pt; color: #00008B;}
body {color: #333333;}
a, a:hover {color: #8B3A62;}
pre {font-size: 12px;}
</style>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(rio,
               tidyverse,
               REAT,
               ConvergenceClubs,
               here)

options(prompt="R> ", digits=4, scipen=999)
```



# Load the Data




```{r load_data, echo=FALSE}
# Goi dataframe
grdp <- import(here("gso", "grdp_vn.xlsx"))


```






# Sigma convergence for two periods (ANOVA): 2018 and 2023 


```{r}
sigma_anova_sd_log <- sigmaconv(
          grdp$y2018,
          2018,
          grdp$y2023,
          2023,
          sigma.measure = "sd",
          sigma.log = TRUE
          ) 
sigma_anova_sd_log

```



# Sigma convergence for multiple periods (Trend regression) 



```{r}
sigma_trend_sd_log <- sigmaconv.t(
   grdp$y2018,
      2018,
   grdp[3:7],
      2023,
 sigma.measure = "sd",
  sigma.log = TRUE) 
sigma_trend_sd_log
  
```



# Linear trend regression


```{r}
sigma_trend_sd_log[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth(method = lm) + 
  theme_minimal() +
  labs(subtitle = "SD[Log(GRDP/nguoi)]",
       x = "",
       y = "")
```



# Non-linear trend regression


```{r}
sigma_trend_sd_log[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth() + 
  theme_minimal() +
  labs(subtitle = "SD[Log(GRDP/nguoi)]",
       x = "",
       y = "")
```


Note: An increasing dispersion implies a lack of sigma convergence




# The analyses of sigma ang beta convergences use trend regression



```{r}
sigma_plot <- rca (grdp$y2018, 2018, grdp[3:7], 2023,
                   conditions = NULL, sigma.type = "trend", sigma.measure = "cv",
                   beta.plot = TRUE, beta.plotLine = TRUE, beta.plotX =
                     "Ln (GRDP/nguoi trung inh)", beta.plotY = "Ln (% tang GRDP/nguoi )",
                   beta.plotTitle = "Hoi tu Beta GRDP/nguoi giai đoan 2018-2023",
                   sigma.plot = TRUE, sigma.plotY = "cv cua ln (GRDP/nguoi)",
                   sigma.plotTitle = "Hội tụ sigma GRDP/nguoi năm 2018-2023")
sigma_plot
#

```



# test if there are average growth differences by mat do doanh nghiep, nsld, FDI, hdi


```{r}

# a. Tao dummy variable
regionaldummies <- to.dummy(grdp$mat_do_dn)
# b. Creating dummy variables for anh huong cua bien so dn/van dan den hoi tu GRDP/nguoi
# regionaldummies[,1] = cao (>=1), 
# regionaldummies[,2] = thap (<1)
grdp$mat_do_dn_van_ld <- regionaldummies[,2]
# # #
converg_grdp <- rca (grdp$y2018, 2018,
                          grdp[3:7], 2023,
                          conditions = grdp[c(16, 19, 20, 26)], # 16, 19, 20, 26 la vi tri cua cacs cot "maty_do_dn_van_ld, ....)
                          sigma.type = "trend")
  converg_grdp

```



# Beta convergence: 2018-2023 period


## OLS


```{r}
betaconv_ols <- betaconv.ols (grdp$y2018,
              2018,
              grdp$y2023,
              2023,
              conditions = NULL, 
              beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid",
              # output.results = TRUE
              )
betaconv_ols

```


### Linear beta-convergence regression 


```{r}
# Create a new column ln_initial = log(y2018)
ln_initial <- grdp %>%
  mutate(ln_initial = log(y2018))

betaconv_ols[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = grdp$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
theme_minimal() +
labs(subtitle = "Growth of grdp (%) (2018-2023)",
   x = "Log of GRDP/nguoi in 2018",
   y = "Growth ò GRDP/nguoi")

```


## NLS


```{r}
betaconv_nls <- betaconv.ols (grdp$y2018,
              2018,
              grdp$y2023,
              2023,
              conditions = NULL, 
              beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid",
              # output.results = TRUE
              )
betaconv_nls

```



```{r}
betaconv_nls[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = grdp$growth)) +
  geom_point() +
  geom_smooth() +
  theme_minimal() +
  labs(subtitle = "Growth of GRDP/nguoi (2018-2023)",
       x = "Log of GRDP/nguoi nam 2018",
       y = "")
```


Note: The lack of statistical significance of the beta coefficient implies a lack of beta convergence




# KẾT LUẬN













END.
