---
title: "Hội tụ năng suất lúa ở ĐBSCL"
author: "Vo Thanh Danh"
date: "2024-09-14"
output:
  html_document:
    code_download: true
    df_print: paged
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
    number_sections: true
    code_folding: "show"
    theme: "cosmo"
    highlight: "monochrome"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_package, echo=FALSE}

pacman::p_load(rio,
               tidyverse,
               REAT,
               ggplot2,
               here,      
               kableExtra,
               tidyverse, 
               here)
```


# Load data md va tg_dist


```{r dbscl, echo=FALSE}

#  Goi/load file: md
md <- import(here("gso", "md.xlsx"))
tkmt_md <- summary(md) 
tkmt_md 

tg_dist <- import(here("gso", "tg_district.xlsx"))
tg_dist %>% 
  select(district, area_total, area_agr, pop, pop_rural) %>% 
  group_by(district) %>% 
  summary()

```


# Load data about ns cac vu dx-dx-ht-td


```{r yiels_conv, echo=FALSE}
# Goi dataframe
ns_annual <- import(here("gso", "tg_ns_annual.xlsx")) # lua ca nam

#
ns_dx <- import(here("gso", "tg_ns_dx.xlsx")) # Dong Xuan

#
ns_ht <- import(here("gso", "tg_ns_ht.xlsx")) # He Thu

#
ns_td <- import(here("gso", "tg_ns_td.xlsx")) # Thu Dong

#
#
# we prevent scientic notation of numbers 
# in R and set a limit of 4 digits:
options(scipen = 100, digits = 4)
#

```




# THỰC TRẠNG SẢN XUẤT LÚA


## Cả năm


```{r lua_ca_nam, echo=FALSE}
# a. tkmt
tg_dist_ann <- tg_dist %>% 
  select(district, annual_rice_area, annual_rice_yield, annual_rice_volumn) %>% 
  group_by(district) %>% 
  summary()
tg_dist_ann

# b. plot
## a. chung 11 huyen
ggplot(data = tg_dist)+
  geom_boxplot(
    mapping = aes(x = year, y = annual_rice_yield, fill = year),
    width = 1, fill = "darkred")+
  theme_minimal()+ # other specification for the geom
  labs(
    x = "Năm",
    y = "Năng suất lúa cả năm- tạ/ha",
    title = "Diễn biến năng suất lúa cả năm giai đoạn 1995-2022")
#
## b. phan theo huyen
ggplot(data = tg_dist)+
  geom_boxplot(
    mapping = aes(x = year, y = annual_rice_yield, fill = year),
    width = 1, fill = "darkred")+
  theme_minimal()+ # other specification for the geom
  labs(
    x = "Năm",
    y = "Năng suất lúa - tạ/ha",
    title = "Diễn biến năng suất lúa cả năm - giai đoạn 1995-2022") +
  facet_wrap(~district)        # Nhom theo tinh

```

## vụ Đông Xuân

```{r dist_dx, echo=FALSE}
# a. tkmt
tg_dist_dx <- tg_dist %>% 
  select(district, win_spr_rice_area, win_spr_rice_yield, win_spr_rice_volumn) %>% 
  group_by(district) %>% 
  summary()
tg_dist_dx

# plot
## a. chung 11 huyen
ggplot(data = tg_dist)+
  geom_boxplot(
    mapping = aes(x = year, y = win_spr_rice_yield, fill = year),
    width = 1, fill = "darkred")+
  theme_minimal()+ # other specification for the geom
  labs(
    x = "Năm",
    y = "Năng suất lúa  vụ Đông Xuân - tạ/ha",
    title = "Diễn biến năng suất lúa vụ Đông Xuân giai đoạn 1995-2022")
#
## b. phan theo huyen
ggplot(data = tg_dist)+
  geom_boxplot(
    mapping = aes(x = year, y = win_spr_rice_yield, fill = year),
    width = 1, fill = "darkred")+
  theme_minimal()+ # other specification for the geom
  labs(
    x = "Năm",
    y = "Năng suất lúa - tạ/ha",
    title = "Diễn biến năng suất lúa vụ Đông Xuân - giai đoạn 1995-2022") +
  facet_wrap(~district)        # Nhom theo tinh

```



## vụ Hè Thu


```{r dist_ht, echo=FALSE}
# a. tkmt
tg_dist_ht <- tg_dist %>% 
  select(district, sum_aut_rice_area, sum_aut_rice_yield, sum_aut_rice_volumn) %>% 
  group_by(district) %>% 
  summary()
tg_dist_ht

# b. plot
## a. chung 11 huyen
ggplot(data = tg_dist)+
  geom_boxplot(
    mapping = aes(x = year, y = sum_aut_rice_yield, fill = year),
    width = 1, fill = "darkred")+
  theme_minimal()+ # other specification for the geom
  labs(
    x = "Năm",
    y = "Năng suất lúa  vụ Hè Thu - tạ/ha",
    title = "Diễn biến năng suất lúa vụ Hè Thu giai đoạn 1995-2022")
#
## b. phan theo huyen
ggplot(data = tg_dist)+
  geom_boxplot(
    mapping = aes(x = year, y = sum_aut_rice_yield, fill = year),
    width = 1, fill = "darkred")+
  theme_minimal()+ # other specification for the geom
  labs(
    x = "Năm",
    y = "Năng suất lúa - tạ/ha",
    title = "Diễn biến năng suất lúa vụ Hè Thu - giai đoạn 1995-2022") +
  facet_wrap(~district)        # Nhom theo tinh

```


## vụ Thu Đông


```{r dist_td, echo=FALSE}
tg_dist_td <- tg_dist %>% 
  select(district, aut_win_rice_area, aut_win_rice_yield, aut_win_rice_volumn) %>% 
  group_by(district) %>% 
  summary()
tg_dist_td

# b. plot
## a. chung 11 huyen
ggplot(data = tg_dist)+
  geom_boxplot(
    mapping = aes(x = year, y = aut_win_rice_yield, fill = year),
    width = 1, fill = "darkred")+
  theme_minimal()+ # other specification for the geom
  labs(
    x = "Năm",
    y = "Năng suất lúa  vụ Thu Đông- tạ/ha",
    title = "Diễn biến năng suất lúa vụ Thu Đông giai đoạn 1995-2022")
#
## b. phan theo huyen
ggplot(data = tg_dist)+
  geom_boxplot(
    mapping = aes(x = year, y = aut_win_rice_yield, fill = year),
    width = 1, fill = "darkred")+
  theme_minimal()+ # other specification for the geom
  labs(
    x = "Năm",
    y = "Năng suất lúa - tạ/ha",
    title = "Diễn biến năng suất lúa vụ Thu Đông - giai đoạn 1995-2022") +
  facet_wrap(~district)        # Nhom theo tinh

```




# PHÂN TÍCH HỘI TỤ NĂNG SUẤT LÚA 



## Cả năm


### test absolute beta convergence comparing the years 2009 and 2019:


#### OLS


```{r annual_yield, echo=FALSE}
# absolute beta convergence comparing the years 2009 and 2019
beta_ols_annual <- betaconv.ols (ns_annual$y2009, 2009, ns_annual$y2019, 2019,
      conditions = NULL, 
      print.results = T,
      beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
      )
beta_ols_annual
#

```




```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- ns_annual %>%
  mutate(ln_initial = log(y2009))

beta_ols_annual[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = ns_annual$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Tang truong nang suat lua ca nam (%) (2009-2019)",
       x = "Log of nang suat luas nam 2009",
       y = "")
```


#### NLS



```{r annual_yield_nls, echo=FALSE}
# absolute beta convergence comparing the years 2009 and 2019
beta_nls_annual <- betaconv.nls (ns_annual$y2009, 2009, ns_annual$y2019, 2019,
      conditions = NULL, 
      print.results = T,
      beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
      )
beta_nls_annual
#

```




```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- ns_annual %>%
  mutate(ln_initial = log(y2009))

beta_nls_annual[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = ns_annual$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Tang truong nang suat lua ca nam (%) (2009-2019)",
       x = "Log of nang suat luas nam 2009",
       y = "")

```





### sigma convergence for two periods (ANOVA): 2009 vs 2019 

```{r}
#
# check/test sigma convergence
sigma_anova_sd_log_annual <- sigmaconv (ns_annual$y2009, 2009, ns_annual$y2019,
                    2019, sigma.measure = "cv")
# Using the coefficient of variation (CV)
# Sigma convergence for two periods (ANOVA)
sigma_anova_sd_log_annual
#

```


### Sigma convergence for multiple periods (Trend regression) 


```{r}
sigma_trend_sd_log_annual <- sigmaconv.t(
   ns_annual$y2009,
      2009,
   ns_annual[22:31],
      2019,
 sigma.measure = "sd",
  sigma.log = TRUE) 
sigma_trend_sd_log_annual
  
```


### Linear trend regression


```{r}
sigma_trend_sd_log_annual[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth(method = lm) + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua ca nam)]",
       x = "",
       y = "")
```



### Non-linear trend regression

```{r}
sigma_trend_sd_log_annual[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth() + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua ca nam)]",
       x = "",
       y = "")
```


Note: An increasing dispersion implies a lack of sigma convergence




### The analysis of sigma convergence uses trend regression


```{r}
sigma_plot <- rca (ns_annual$y2009, 2009, ns_annual[22:31], 2019,
                   conditions = NULL, sigma.type = "trend", sigma.measure = "cv",
                   beta.plot = TRUE, beta.plotLine = TRUE, beta.plotX =
                     "Ln (Nang suat trung binh)", beta.plotY = "Ln (Tang nang suat trung binh)",
                   beta.plotTitle = "Hoi tu Beta nang suat lua ca nam giai đoan 2009-2019",
                   sigma.plot = TRUE, sigma.plotY = "cv cua ln (Nang suat lua)",
                   sigma.plotTitle = "Hội tụ sigma năng suất lúa cả năm 2009-2019")
sigma_plot
#

```



### test if there are average growth differences between districts


```{r}
#
# a. Tao dummy variable
regionaldummies <- to.dummy(ns_annual$coastal)
# b. Creating dummy variables for giáp biến vs. không giáp biển
# regionaldummies[,1] = Giáp biển (1/0), regionaldummies[,2] = không giáp biển (1/0)
ns_annual$coastal <- regionaldummies[,2]
# 
converg_annual <- rca (ns_annual$y2009, 2009,
                        ns_annual[22:31], 2019, 
                        conditions = ns_annual[c(36)],
                        sigma.type = "trend")
converg_annual

```



##  vụ Đông Xuân


### test absolute beta convergence comparing the years 2009 and 2019:


#### OLS


```{r dx_yield, echo=FALSE}
# absolute beta convergence comparing the years 2009 and 2019
beta_ols_dx <- betaconv.ols (ns_dx$y2009, 2009, ns_dx$y2019, 2019,
      conditions = NULL, 
      print.results = T,
      beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
      )
beta_ols_dx
#

```




```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- ns_dx %>%
  mutate(ln_initial = log(y2009))

beta_ols_dx[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = ns_dx$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Tang truong nang suat lua ca nam (%) (2009-2019)",
       x = "Log of nang suat luas nam 2009",
       y = "")
```


#### NLS



```{r dx_yield_nls, echo=FALSE}
# absolute beta convergence comparing the years 2009 and 2019
beta_nls_dx <- betaconv.nls (ns_dx$y2009, 2009, ns_dx$y2019, 2019,
      conditions = NULL, 
      print.results = T,
      beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
      )
beta_nls_dx
#

```




```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- ns_dx %>%
  mutate(ln_initial = log(y2009))

beta_nls_dx[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = ns_dx$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Tang truong nang suat lua ca nam (%) (2009-2019)",
       x = "Log of nang suat luas nam 2009",
       y = "")

```





### sigma convergence for two periods (ANOVA): 2009 vs 2019 

```{r}
#
# check/test sigma convergence
sigma_anova_sd_log_dx <- sigmaconv (ns_dx$y2009, 2009, ns_dx$y2019,
                    2019, sigma.measure = "cv")
# Using the coefficient of variation (CV)
# Sigma convergence for two periods (ANOVA)
sigma_anova_sd_log_dx
#

```


### Sigma convergence for multiple periods (Trend regression) 


```{r}
sigma_trend_sd_log_dx <- sigmaconv.t(
   ns_dx$y2009,
      2009,
   ns_dx[22:31],
      2019,
 sigma.measure = "sd",
  sigma.log = TRUE) 
sigma_trend_sd_log_dx
  
```


### Linear trend regression


```{r}
sigma_trend_sd_log_dx[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth(method = lm) + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua ca nam)]",
       x = "",
       y = "")
```



### Non-linear trend regression

```{r}
sigma_trend_sd_log_dx[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth() + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua ca nam)]",
       x = "",
       y = "")
```


Note: An increasing dispersion implies a lack of sigma convergence




### The analysis of sigma convergence uses trend regression


```{r}
sigma_plot <- rca (ns_dx$y2009, 2009, ns_dx[22:31], 2019,
                   conditions = NULL, sigma.type = "trend", sigma.measure = "cv",
                   beta.plot = TRUE, beta.plotLine = TRUE, beta.plotX =
                     "Ln (Nang suat trung binh)", beta.plotY = "Ln (Tang nang suat trung binh)",
                   beta.plotTitle = "Hoi tu Beta nang suat lua ca nam giai đoan 2009-2019",
                   sigma.plot = TRUE, sigma.plotY = "cv cua ln (Nang suat lua)",
                   sigma.plotTitle = "Hội tụ sigma năng suất lúa cả năm 2009-2019")
sigma_plot
#

```



### test if there are average growth differences between districts


```{r}
#
# a. Tao dummy variable
regionaldummies <- to.dummy(ns_dx$coastal)
# b. Creating dummy variables for giáp biến vs. không giáp biển
# regionaldummies[,1] = Giáp biển (1/0), regionaldummies[,2] = không giáp biển (1/0)
ns_dx$coastal <- regionaldummies[,2]
# 
converg_dx <- rca (ns_dx$y2009, 2009,
                        ns_dx[22:31], 2019, 
                        conditions = ns_dx[c(36)],
                        sigma.type = "trend")
converg_dx

```




## Vụ Hè Thu


### test absolute beta convergence comparing the years 2009 and 2019:


#### OLS


```{r ht_yield, echo=FALSE}
# absolute beta convergence comparing the years 2009 and 2019
beta_ols_ht <- betaconv.ols (ns_ht$y2009, 2009, ns_ht$y2019, 2019,
      conditions = NULL, 
      print.results = T,
      beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
      )
beta_ols_ht
#

```




```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- ns_ht %>%
  mutate(ln_initial = log(y2009))

beta_ols_ht[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = ns_ht$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Tang truong nang suat lua ca nam (%) (2009-2019)",
       x = "Log of nang suat luas nam 2009",
       y = "")
```


#### NLS



```{r ht_yield_nls, echo=FALSE}
# absolute beta convergence comparing the years 2009 and 2019
beta_nls_ht <- betaconv.nls (ns_ht$y2009, 2009, ns_ht$y2019, 2019,
      conditions = NULL, 
      print.results = T,
      beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
      )
beta_nls_ht
#

```




```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- ns_ht %>%
  mutate(ln_initial = log(y2009))

beta_nls_ht[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = ns_ht$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Tang truong nang suat lua ca nam (%) (2009-2019)",
       x = "Log of nang suat luas nam 2009",
       y = "")

```





### sigma convergence for two periods (ANOVA): 2009 vs 2019 

```{r}
#
# check/test sigma convergence
sigma_anova_sd_log_ht <- sigmaconv (ns_ht$y2009, 2009, ns_ht$y2019,
                    2019, sigma.measure = "cv")
# Using the coefficient of variation (CV)
# Sigma convergence for two periods (ANOVA)
sigma_anova_sd_log_ht
#

```


### Sigma convergence for multiple periods (Trend regression) 


```{r}
sigma_trend_sd_log_ht <- sigmaconv.t(
   ns_ht$y2009,
      2009,
   ns_ht[22:31],
      2019,
 sigma.measure = "sd",
  sigma.log = TRUE) 
sigma_trend_sd_log_ht
  
```


### Linear trend regression


```{r}
sigma_trend_sd_log_ht[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth(method = lm) + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua ca nam)]",
       x = "",
       y = "")
```



### Non-linear trend regression

```{r}
sigma_trend_sd_log_ht[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth() + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua ca nam)]",
       x = "",
       y = "")
```


Note: An increasing dispersion implies a lack of sigma convergence




### The analysis of sigma convergence uses trend regression


```{r}
sigma_plot <- rca (ns_ht$y2009, 2009, ns_ht[22:31], 2019,
                   conditions = NULL, sigma.type = "trend", sigma.measure = "cv",
                   beta.plot = TRUE, beta.plotLine = TRUE, beta.plotX =
                     "Ln (Nang suat trung binh)", beta.plotY = "Ln (Tang nang suat trung binh)",
                   beta.plotTitle = "Hoi tu Beta nang suat lua ca nam giai đoan 2009-2019",
                   sigma.plot = TRUE, sigma.plotY = "cv cua ln (Nang suat lua)",
                   sigma.plotTitle = "Hội tụ sigma năng suất lúa cả năm 2009-2019")
sigma_plot
#

```



### test if there are average growth differences between districts


```{r}
#
# a. Tao dummy variable
regionaldummies <- to.dummy(ns_ht$coastal)
# b. Creating dummy variables for giáp biến vs. không giáp biển
# regionaldummies[,1] = Giáp biển (1/0), regionaldummies[,2] = không giáp biển (1/0)
ns_ht$coastal <- regionaldummies[,2]
# 
converg_ht <- rca (ns_ht$y2009, 2009,
                        ns_ht[22:31], 2019, 
                        conditions = ns_ht[c(36)],
                        sigma.type = "trend")
converg_ht

```




## Vụ Thu Đông


### test absolute beta convergence comparing the years 2009 and 2019:


#### OLS


```{r td_yield, echo=FALSE}
# absolute beta convergence comparing the years 2009 and 2019
beta_ols_td <- betaconv.ols (ns_td$y2009, 2009, ns_td$y2019, 2019,
      conditions = NULL, 
      print.results = T,
      beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
      )
beta_ols_td
#

```




```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- ns_td %>%
  mutate(ln_initial = log(y2009))

beta_ols_td[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = ns_td$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Tang truong nang suat lua ca nam (%) (2009-2019)",
       x = "Log of nang suat luas nam 2009",
       y = "")
```


#### NLS



```{r td_yield_nls, echo=FALSE}
# absolute beta convergence comparing the years 2009 and 2019
beta_nls_td <- betaconv.nls (ns_td$y2009, 2009, ns_td$y2019, 2019,
      conditions = NULL, 
      print.results = T,
      beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
      )
beta_nls_td
#

```




```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- ns_td %>%
  mutate(ln_initial = log(y2009))

beta_nls_td[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = ns_td$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Tang truong nang suat lua ca nam (%) (2009-2019)",
       x = "Log of nang suat luas nam 2009",
       y = "")

```





### sigma convergence for two periods (ANOVA): 2009 vs 2019 

```{r}
#
# check/test sigma convergence
sigma_anova_sd_log_td <- sigmaconv (ns_td$y2009, 2009, ns_td$y2019,
                    2019, sigma.measure = "cv")
# Using the coefficient of variation (CV)
# Sigma convergence for two periods (ANOVA)
sigma_anova_sd_log_td
#

```


### Sigma convergence for multiple periods (Trend regression) 


```{r}
sigma_trend_sd_log_td <- sigmaconv.t(
   ns_td$y2009,
      2009,
   ns_td[22:31],
      2019,
 sigma.measure = "sd",
  sigma.log = TRUE) 
sigma_trend_sd_log_td
  
```


### Linear trend regression


```{r}
sigma_trend_sd_log_td[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth(method = lm) + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua ca nam)]",
       x = "",
       y = "")
```



### Non-linear trend regression

```{r}
sigma_trend_sd_log_td[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth() + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua ca nam)]",
       x = "",
       y = "")
```


Note: An increasing dispersion implies a lack of sigma convergence




### The analysis of sigma convergence uses trend regression


```{r}
sigma_plot <- rca (ns_td$y2009, 2009, ns_td[22:31], 2019,
                   conditions = NULL, sigma.type = "trend", sigma.measure = "cv",
                   beta.plot = TRUE, beta.plotLine = TRUE, beta.plotX =
                     "Ln (Nang suat trung binh)", beta.plotY = "Ln (Tang nang suat trung binh)",
                   beta.plotTitle = "Hoi tu Beta nang suat lua ca nam giai đoan 2009-2019",
                   sigma.plot = TRUE, sigma.plotY = "cv cua ln (Nang suat lua)",
                   sigma.plotTitle = "Hội tụ sigma năng suất lúa cả năm 2009-2019")
sigma_plot
#

```



### test if there are average growth differences between districts


```{r}
#
# a. Tao dummy variable
regionaldummies <- to.dummy(ns_td$coastal)
# b. Creating dummy variables for giáp biến vs. không giáp biển
# regionaldummies[,1] = Giáp biển (1/0), regionaldummies[,2] = không giáp biển (1/0)
ns_td$coastal <- regionaldummies[,2]
# 
converg_td <- rca (ns_td$y2009, 2009,
                        ns_td[22:31], 2019, 
                        conditions = ns_td[c(36)],
                        sigma.type = "trend")
converg_td

```







END.








