---
title: "Classical Sigma and Beta Convergence Analysis of rice yield for Mekong Delta:"
subtitle: "Dong Xuan and He Thu-Thu Dong"
author: "Vo Thanh Danh"
output:
  html_document:
    code_download: true
    df_print: paged
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
    number_sections: true
    code_folding: "show"
    theme: "cosmo"
    highlight: "monochrome"
  pdf_document: default
  word_document: default
---


<style>
h1.title {font-size: 18pt; color: DarkBlue;} 
body, h1, h2, h3, h4 {font-family: "Palatino", serif;}
body {font-size: 12pt;}
/* Headers */
h1,h2,h3,h4,h5,h6{font-size: 14pt; color: #00008B;}
body {color: #333333;}
a, a:hover {color: #8B3A62;}
pre {font-size: 12px;}
</style>




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(rio,
               tidyverse,
               REAT,
               ConvergenceClubs,
               here)

options(prompt="R> ", digits=4, scipen=999)
```



# Load the Data



```{r load_data, echo=FALSE}
# Goi dataframe
md_dx <- import(here("weiland", "gso", "ns_dx_1.xlsx"))
md_ht <- import(here("weiland", "gso", "ns_ht_1.xlsx"))

```



# Vu Dong Xuan


## Sigma convergence for two periods (ANOVA): 1995 and 2023 


```{r}
sigma_anova_sd_log_dx <- sigmaconv(
          md_dx$y1995,
          1995,
          md_dx$y2023,
          2023,
          sigma.measure = "sd",
          sigma.log = TRUE) 
sigma_anova_sd_log_dx

```



## Sigma convergence for multiple periods (Trend regression) 



```{r}
sigma_trend_sd_log_dx <- sigmaconv.t(
   md_dx$y1995,
      1995,
   md_dx[3:30],
      2023,
 sigma.measure = "sd",
  sigma.log = TRUE) 
sigma_trend_sd_log_dx
  
```



## Linear trend regression


```{r}
sigma_trend_sd_log_dx[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth(method = lm) + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua vu Dong Xuan)]",
       x = "",
       y = "")
```



## Non-linear trend regression

```{r}
sigma_trend_sd_log_dx[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth() + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua vu Dong Xuan)]",
       x = "",
       y = "")
```


Note: An increasing dispersion implies a lack of sigma convergence





## The analysis of sigma convergence uses trend regression


```{r}
sigma_plot <- rca (md_dx$y1995, 1995, md_dx[3:30], 2023,
                   conditions = NULL, sigma.type = "trend", sigma.measure = "cv",
                   beta.plot = TRUE, beta.plotLine = TRUE, beta.plotX =
                     "Ln (Nang suat trung binh)", beta.plotY = "Ln (Tang nang suat trung binh)",
                   beta.plotTitle = "Hoi tu Beta nang suat lua vu Dong Xuan giai đoan 2009-2019",
                   sigma.plot = TRUE, sigma.plotY = "cv cua ln (Nang suat lua)",
                   sigma.plotTitle = "Hội tụ sigma năng suất lúa cả năm 2009-2019")
sigma_plot
#

```



## test if there are average growth differences between provinces


```{r}
#
# a. Tao dummy variable
regionaldummies <- to.dummy(md_dx$Province)
# b. Creating dummy variables for giáp biến vs. không giáp biển
# regionaldummies[,1] = Giáp biển (1/0), regionaldummies[,2] = không giáp biển (1/0)
md_dx$Province <- regionaldummies[,2]
# 
converg_dx <- rca (md_dx$y1995, 1995,
                        md_dx[3:30], 2023, 
                        conditions = md_dx[c(31, 32)],
                        sigma.type = "trend")
converg_dx

```



## Beta convergence: 1995-2023 period


### OLS


```{r}
betaconv_ols_dx <- betaconv.ols (md_dx$y1995,
              1995,
              md_dx$y2023,
              2023,
              conditions = NULL, 
              beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid",
              # output.results = TRUE
              )
betaconv_ols_dx

```


 Linear beta-convergence regression 


```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- md_dx %>%
  mutate(ln_initial = log(y1995))

betaconv_ols_dx[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = md_dx$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Growth of rice yield (1995-2023)",
       x = "Log of rice yield in 1995",
       y = "")
```


### NLS


```{r}
betaconv_nls_dx <- betaconv.ols (md_dx$y1995,
              1995,
              md_dx$y2023,
              2023,
              conditions = NULL, 
              beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid",
              # output.results = TRUE
              )
betaconv_nls_dx

```



```{r}
betaconv_nls_dx[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = md_dx$growth)) +
  geom_point() +
  geom_smooth() +
  theme_minimal() +
  labs(subtitle = "Growth of rice yield (1995-2023)",
       x = "Log of rice yield in 1995",
       y = "")
```


Note: The lack of statistical significance of the beta coefficient implies a lack of beta convergence







# Vu He Thu - Thu Dong


## Sigma convergence for two periods (ANOVA): 1995 and 2023 


```{r}
sigma_anova_sd_log_ht <- sigmaconv(
          md_ht$y1995,
          1995,
          md_ht$y2023,
          2023,
          sigma.measure = "sd",
          sigma.log = TRUE)
sigma_anova_sd_log_ht          
```



## Sigma convergence for multiple periods (Trend regression) 
 

```{r}
sigma_trend_sd_log_ht <- sigmaconv.t(
   md_ht$y1995,
      1995,
   md_ht[3:30],
      2023,
 sigma.measure = "sd",
  sigma.log = TRUE) 
sigma_trend_sd_log_ht
  
```

## Linear trend regression

```{r}
sigma_trend_sd_log_ht[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth(method = lm) + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua vu Dong Xuan)]",
       x = "",
       y = "")
```

## Non-linear trend regression

```{r}
sigma_trend_sd_log_ht[["sigma.trend"]] %>% 
  ggplot(aes(x = years, y = sigma.years)) +
  geom_line(size=1, linetype = "dashed") + 
  geom_smooth() + 
  theme_minimal() +
  labs(subtitle = "SD[Log(Nang suat lua vu Dong Xuan)]",
       x = "",
       y = "")
```


Note: An increasing dispersion implies a lack of sigma convergence




## The analysis of sigma convergence uses trend regression


```{r}
sigma_plot <- rca (md_ht$y1995, 1995, md_ht[3:30], 2023,
                   conditions = NULL, sigma.type = "trend", sigma.measure = "cv",
                   beta.plot = TRUE, beta.plotLine = TRUE, beta.plotX =
                     "Ln (Nang suat trung binh)", beta.plotY = "Ln (Tang nang suat trung binh)",
                   beta.plotTitle = "Hoi tu Beta nang suat lua vu Dong Xuan giai đoan 2009-2019",
                   sigma.plot = TRUE, sigma.plotY = "cv cua ln (Nang suat lua)",
                   sigma.plotTitle = "Hội tụ sigma năng suất lúa cả năm 2009-2019")
sigma_plot
#

```



## test if there are average growth differences between provinces


```{r}
#
# a. Tao dummy variable
regionaldummies <- to.dummy(md_ht$Province)
# b. Creating dummy variables for giáp biến vs. không giáp biển
# regionaldummies[,1] = Giáp biển (1/0), regionaldummies[,2] = không giáp biển (1/0)
md_ht$Province <- regionaldummies[,2]
# 
converg_ht <- rca (md_ht$y1995, 1995,
                        md_ht[3:30], 2023, 
                        conditions = md_ht[c(31, 32)],
                        sigma.type = "trend")
converg_ht

```





## Beta convergence: 1995-2023 period


### OLS


```{r}
betaconv_ols_ht <- betaconv.ols (md_ht$y1995,
              1995,
              md_ht$y2023,
              2023,
              conditions = NULL, 
              beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
              )
betaconv_ols_ht
```



```{r}
# Create a new column ln_initial = log(y1995)
ln_initial <- md_ht %>%
  mutate(ln_initial = log(y1995))

betaconv_ols_ht[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = md_ht$growth)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  labs(subtitle = "Growth of rice yield (1995-2023)",
       x = "Log of rice yield in 1995",
       y = "")
```



### NLS



```{r}
betaconv_nls_ht <- betaconv.ols (md_ht$y1995,
              1995,
              md_ht$y2023,
              2023,
              conditions = NULL, 
              beta.plot = TRUE,
              beta.plotLine = TRUE, 
              beta.plotLineCol = "red", 
              beta.plotX = "ln (initial)", 
              beta.plotY = "growth", 
              beta.plotTitle = "Beta convergence", 
              beta.bgrid = TRUE, 
              beta.bgridType = "solid"
              )
betaconv_nls_ht
```



```{r}
betaconv_nls_ht[["regdata"]] %>%
  ggplot(aes(x = ln_initial, y = md_ht$growth)) +
  geom_point() +
  geom_smooth() +
  theme_minimal() +
  labs(subtitle = "Growth of rice yield (1995-2023)",
       x = "Log of rice yield in 1995",
       y = "")
```


Note: The lack of statistical significance of the beta coefficient implies a lack of beta convergence





END.
